### 2019-03-12 paper/images/handbags_edges_lotsofresults_latex/fig.tex

\begin{figure*}
\begin{center}
\bgroup 
 \def\arraystretch{0.2} 
 \setlength\tabcolsep{0.2pt}
\begin{tabular}{cccccc}
Input & Ground truth & Output & Input & Ground truth & Output \\ 
%\includegraphics[width=0.16666666666667\linewidth]{figs/handbags_edges_lotsofresults_latex/input_128_AB.jpg} &
%\includegraphics[width=0.16666666666667\linewidth]{figs/handbags_edges_lotsofresults_latex/gt_128_AB.jpg} &
%\includegraphics[width=0.16666666666667\linewidth]{figs/handbags_edges_lotsofresults_latex/L1cGAN_128_AB.jpg} \hspace{0.025in} &
%\includegraphics[width=0.16666666666667\linewidth]{figs/handbags_edges_lotsofresults_latex/input_146_AB.jpg} &
%\includegraphics[width=0.16666666666667\linewidth]{figs/handbags_edges_lotsofresults_latex/gt_146_AB.jpg} &
%\includegraphics[width=0.16666666666667\linewidth]{figs/handbags_edges_lotsofresults_latex/L1cGAN_146_AB.jpg} \\ 
\includegraphics[width=0.16666666666667\linewidth]{figs/handbags_edges_lotsofresults_latex/input_12_AB.jpg} &
\includegraphics[width=0.16666666666667\linewidth]{figs/handbags_edges_lotsofresults_latex/gt_12_AB.jpg} &
\includegraphics[width=0.16666666666667\linewidth]{figs/handbags_edges_lotsofresults_latex/L1cGAN_12_AB.jpg} \hspace{0.025in} &
\includegraphics[width=0.16666666666667\linewidth]{figs/handbags_edges_lotsofresults_latex/input_157_AB.jpg} &
\includegraphics[width=0.16666666666667\linewidth]{figs/handbags_edges_lotsofresults_latex/gt_157_AB.jpg} &
\includegraphics[width=0.16666666666667\linewidth]{figs/handbags_edges_lotsofresults_latex/L1cGAN_157_AB.jpg} \\ 
\includegraphics[width=0.16666666666667\linewidth]{figs/handbags_edges_lotsofresults_latex/input_130_AB.jpg} &
\includegraphics[width=0.16666666666667\linewidth]{figs/handbags_edges_lotsofresults_latex/gt_130_AB.jpg} &
\includegraphics[width=0.16666666666667\linewidth]{figs/handbags_edges_lotsofresults_latex/L1cGAN_130_AB.jpg} \hspace{0.025in} &
\includegraphics[width=0.16666666666667\linewidth]{figs/handbags_edges_lotsofresults_latex/input_119_AB.jpg} &
\includegraphics[width=0.16666666666667\linewidth]{figs/handbags_edges_lotsofresults_latex/gt_119_AB.jpg} &
\includegraphics[width=0.16666666666667\linewidth]{figs/handbags_edges_lotsofresults_latex/L1cGAN_119_AB.jpg} \\ 
\includegraphics[width=0.16666666666667\linewidth]{figs/handbags_edges_lotsofresults_latex/input_62_AB.jpg} &
\includegraphics[width=0.16666666666667\linewidth]{figs/handbags_edges_lotsofresults_latex/gt_62_AB.jpg} &
\includegraphics[width=0.16666666666667\linewidth]{figs/handbags_edges_lotsofresults_latex/L1cGAN_62_AB.jpg} \hspace{0.025in} &
\includegraphics[width=0.16666666666667\linewidth]{figs/handbags_edges_lotsofresults_latex/input_106_AB.jpg} &
\includegraphics[width=0.16666666666667\linewidth]{figs/handbags_edges_lotsofresults_latex/gt_106_AB.jpg} &
\includegraphics[width=0.16666666666667\linewidth]{figs/handbags_edges_lotsofresults_latex/L1cGAN_106_AB.jpg}

\end{tabular} \egroup 
\end{center}
\caption{Example results of our method on automatically detected edges$\rightarrow$handbags, compared to ground truth.}
\label{handbags_edges_lotsofresults}
\end{figure*}