### 2019-03-13 paper/presentation.tex
Drafted some ideas for next week's presentation

\documentclass{beamer}
\usetheme{Rochester}

\title[Composing like a human]{Composing like a human}
\subtitle{Adapting generative networks to few-shot learning in the musical domain}
\author{Tudor Paisa  \texorpdfstring{\\ (SNR: 2019551, ANR: 315146)} \and \texttt{t.paisa@tilburguniversity.edu} }
\institute{Tilburg University}
\date{\today}

\begin{document}
\begin{frame}
    \titlepage
\end{frame}

\begin{frame}
    \frametitle{Outline}
    \tableofcontents
\end{frame}

\begin{frame}
    \frametitle{Introduction}
    \section{Introduction}
    \subsection{Deep Learning}
    \begin{itemize}
        \item Deep Learning promises to discover rich hierarchical models over various applications
    \end{itemize}

\end{frame}

\end{document}