### 2019-05-10 paper/images/c_rnn_gan/c_rnn_gan_1_5beethoven-transposed.tex

% This file was created by matplotlib2tikz v0.7.4.
\begin{tikzpicture}

\begin{axis}[
height=\heatheight,
tick align=outside,
tick pos=left,
width=\heatwidth,
x grid style={white!69.01960784313725!black},
xlabel={Events Sequences},
xmin=-0.5, xmax=199.5,
xtick style={color=black},
y grid style={white!69.01960784313725!black},
ylabel={Activations},
ymin=-0.5, ymax=349.5,
ytick style={color=black}
]
\addplot graphics [includegraphics cmd=\pgfimage,xmin=-0.5, xmax=199.5, ymin=349.5, ymax=-0.5] {images/c_rnn_gan/c_rnn_gan_1_5beethoven-transposed-000.png};
\end{axis}

\end{tikzpicture}