### 2019-05-10 paper/images/performance_rnn/performance_rnn_1_5-gen-.tex

% This file was created by matplotlib2tikz v0.7.4.
\begin{tikzpicture}

\begin{axis}[
height=\heatheight,
tick align=outside,
tick pos=left,
width=\heatwidth,
x grid style={white!69.01960784313725!black},
xlabel={Vocabulary},
xmin=-0.5, xmax=415.5,
xtick style={color=black},
y grid style={white!69.01960784313725!black},
ylabel={Event Sequences},
ymin=-0.5, ymax=199.5,
ytick style={color=black}
]
\addplot graphics [includegraphics cmd=\pgfimage,xmin=-0.5, xmax=415.5, ymin=199.5, ymax=-0.5] {images/performance_rnn/performance_rnn_1_5-gen--001.png};
\end{axis}

\end{tikzpicture}